whcB---
title: "Working-Hotelling Bands for Simple LInear Regression"
author: "Laura Bernhofen"
date: "3/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Using the loess smoother to check linearity - see Sec 3.10 in the text *Applied Linear Regression Models*, 4th ed.

1.  Create confidence bands for the model $E(Y) = \beta_0 + \beta_1 X$

  * Note:  You can choose the confidence level by changing the value of "level = " in the first line of the code.  I'll use level = 0.95. 

```{r whbands}
#' Working-Hotelling bands for simple linear regression.
#'
#' Intervals of the form "fit +/- w * standard-error", where w^2 is 
#' found by \code{p * qf(level, p, n - p)}.
#'
#' @param object An object of class "lm".
#' @param newdata A data frame containing the new data.
#' @param level The confidence level of the band.
#'
#' @author David Gerard
whbands <- function(object, newdata, level = 0.95) {
  stopifnot(inherits(object, "lm"))
  stopifnot(inherits(newdata, "data.frame"))
  stopifnot(is.numeric(level), length(level) == 1)
  pout <- stats::predict(object = object, 
                         newdata = newdata, 
                         se.fit = TRUE, 
                         interval = "none")
  n <- nrow(stats::model.matrix(object))
  p <- ncol(stats::model.matrix(object))
  w <- sqrt(p * stats::qf(p = level, df1 = p, df2 = n - p))
  lwr <- pout$fit - w * pout$se.fit
  upr <- pout$fit + w * pout$se.fit
  pout$fit <- cbind(fit = pout$fit, lwr = lwr, upr = upr)
  return(pout)
}
```


Apply the **whbands(object, newdata)** function which we defined in the code chunk above to our model to get the 95% confidence bands for $E(Y)$ and add the smoothing function loess to our plot

```{r}
library(tidyverse)
```


```{r conf-bands-loess,  fig.dim=c(5,3)}
rmr <- read_csv("rmr.csv")
mrout <- lm(metabolic.rate ~ body.weight, data = rmr)
newdf <- data.frame(body.weight = seq(from = min(rmr$body.weight),
                                      to = max(rmr$body.weight), length.out = 100))


whfit <- whbands(object = mrout, newdata = newdf) 
whfit$fit %>%
  cbind(newdf) -> 
  newdf

ggplot() +
  geom_point(data = rmr, mapping = aes(x = body.weight, y = metabolic.rate)) +
  geom_smooth(data = rmr, mapping = aes(x = body.weight, y = metabolic.rate), method = 'lm', se = FALSE) +
  geom_line(data = newdf, mapping = aes(x = body.weight, y = lwr)) +
  geom_line(data = newdf, mapping = aes(x = body.weight, y = upr)) +
  geom_smooth(data = rmr, mapping = aes(x = body.weight, y = metabolic.rate), method = 'loess', color = "orange", se = FALSE)

```

* Linear Model does a good job of fitting our data if the loess (locally weighted regression scatter plot smoothing)  remains within the confidence bands of $E(Y)$
